\chapter{Related Publications}
Different parts of the work presented in this chapter are published in the form of articles in international conferences and workshops.
Parts of this chapter are also published in the form of internal technical reports.
\begin{itemize}
\item \emph{Infrastructure for data processing in large-scale interconnected sensor networks}, Karl Aberer , Manfred Hauswirth , Ali Salehi. Mobile Data Management (MDM), Germany, 2007.
\item \emph{GSN, Quick and Simple Sensor Network Deployment}, Ali Salehi, Karl Aberer. European conference on Wireless Sensor Networks (EWSN), Netherlands, 2007.
\item \emph{Zero-programming Sensor Network Deployment}, Karl Aberer , Manfred Hauswirth , Ali Salehi. Next Generation Service Platforms for Future Mobile Systems (SPMS), Japan, 2007.
\item \emph{A middleware for fast and flexible sensor network deployment}, Karl Aberer , Manfred Hauswirth , Ali Salehi. Very Large Data Bases (VLDB) Seoul, Korea, 2006.
\item \emph{Middleware support for the "Internet of Things"}, Karl Aberer , Manfred Hauswirth , Ali Salehi. 5. GI/ITG KuVS Fachgesprch "Drahtlose Sensornetze", Universitt Stuttgart, 2006.
\item \emph{The Global Sensor Networks middleware for efficient and flexible deployment and interconnection of sensor networks}, Karl Aberer , Manfred Hauswirth , Ali Salehi.
Technical Report, LSIR-2006-006.
\item \emph{Global Sensor Networks}, Karl Aberer , Manfred Hauswirth , Ali Salehi. Technical Report, LSIR-2006-001.
\end{itemize}
\newpage